Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-05-11 Moltres.tex

Molten salt reactor concepts garnered considerable interest in the 1950s and
60s with development of the \gls{ARE} and later the \gls{MSRE} at \gls{ORNL}.
With the inclusion of the \gls{MSR} among the Generation-IV reactor designs
\cite{gen_IV}, this reactor concept has gained renewed research interest in the
past decade, with many new nuclear companies proposing both fluid-fuelled and
solid-fuelled commerical \gls{MSR} designs
\cite{terrapower,terrestrial,elysium,thorcon,kairos}. Two key advantages
offered by the fluid-fuelled \gls{MSR} are improved fuel utilization and no
radiation damage constraint on attainable fuel burn-up. Together, these
attributes result in significantly reduced core fissile inventories and spent
nuclear fuel mass \cite{noauthor_gen_2008}.  These characteristics give the
\gls{MSR} a potentially unique capability for actinide burning and extension of
fuel resources.

@@

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_arfc/moltres_nodate}, for \gls{MSR} simulation.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units. Moltres is devoted to previously unmatched
fidelity in coupled neutronics and thermal hydraulics \gls{MSR} simulation.

Moltres is an open source code licensed under \gls{LGPL} terms that aims to be a tool that can be
freely used, studied, interrogated, and improved by the \gls{MSR}
community.  Its openness is a defining characteristic and
promotes unmatched quality through transparency and ease of peer review. In the
era of GitHub\cite{github} and international scientific collaboration, open and modern software practices must be
employed in order for nuclear engineering simulation capability to enable
discovery and support the regulatory needs presented by new reactor designs.
Furthermore, to openness is necessary to keep pace with scientific computing
advances in fields like astronomy whose multitude of open simulation tools
(Enzo\cite{enzo}, Gadget\cite{gadget}, Flash\cite{flash}, Gizmo\cite{gizmo},
Ramses\cite{ramses}, Mesa\cite{mesa}) allows them to be the darlings of
scientific visualization and public interest.

In that vein, Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another \gls{LGPL}
code that itself leans on LibMesh \cite{kirk2006libmesh}, a \gls{LGPL} finite element
library, and PetSc \cite{petsc-user-ref}, a \gls{BSD}-licensed toolkit for solving nonlinear
equations stemming from discretization of PDEs. \gls{MOOSE} and LibMesh handle
translation of application (Moltres) defined weak PDE forms into residual and
Jacobian functions that are the inputs into PetSc Newton-Raphson solution
routines. All codes use MPI for parallel communication and are easily deployed
on massively-parallel cluster-computing platforms. \gls{MOOSE} applications by default
use monolithic and implicit methods that are ideal for closely-coupled and
multi-scale physics. These methods are ideal for the model problem described in
this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it very
extensible to myriad future modeling challenges.

>>>>

Molten salt reactor concepts garnered considerable interest in the 1950s and 60s
with development of the \gls{ARE} and later the \gls{MSRE} at \gls{ORNL}.  With
the inclusion of the \gls{MSR} among the Generation-IV reactor designs
\cite{noauthor_gif_2015}, this reactor concept has gained renewed research
interest in the past decade, with many new nuclear companies proposing both
fluid-fuelled and solid-fuelled commerical \gls{MSR} designs
\cite{hyde_liquid_2015,leblanc_integral_2015,noauthor_-able_2017,scarlat_design_2014,noauthor_neutronics_2016}. Two
key advantages offered by the fluid-fuelled \gls{MSR} are improved fuel
utilization and no radiation damage constraint on attainable fuel
burn-up. Together, these attributes result in significantly reduced core fissile
inventories and spent nuclear fuel mass \cite{noauthor_gen_2008}.  These
characteristics give the \gls{MSR} a potentially unique capability for actinide
burning and extension of fuel resources.

@@

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_arfc/moltres_nodate}, for \gls{MSR} simulation.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units. Through its computing power, Moltres is devoted to previously
unmatched fidelity in coupled neutronics and thermal hydraulics \gls{MSR}
simulation.

Moltres is an open source code licensed under \gls{LGPL} terms that aims to be a
tool that can be freely used, studied, interrogated, and improved by the
\gls{MSR} community.  Its openness is a defining characteristic and promotes
quality through transparency and ease of peer review. In the era of
GitHub\cite{noauthor_build_nodate} and international scientific collaboration,
open and modern software practices must be employed in order for nuclear
engineering simulation capability to enable discovery and support the regulatory
needs presented by new reactor designs. In that vein, Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another
\gls{LGPL} code that itself leans on LibMesh \cite{kirk_libmesh:_2006}, a
\gls{LGPL} finite element library, and PetSc \cite{satish_balay_petsc_2015}, a
\gls{BSD}-licensed toolkit for solving nonlinear equations stemming from
discretization of PDEs. \gls{MOOSE} and LibMesh handle translation of
application (Moltres) defined weak PDE forms into residual and Jacobian
functions that are the inputs into PetSc Newton-Raphson solution routines. All
codes use MPI for parallel communication and are easily deployed on
massively-parallel cluster-computing platforms. \gls{MOOSE} applications by
default use monolithic and implicit methods that are ideal for closely-coupled
and multi-scale physics. These methods are ideal for the model problem described
in this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it very
extensible to myriad future modeling challenges.
